### 2017-03-06 tssim.tex
a few tweaks

\documentclass{bioinfo}
\copyrightyear{2017} \pubyear{2017}

\newcommand{\plr}[1]{{\color{blue}\it #1}}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Application Notes}

\begin{document}
\firstpage{1}

% Bioinformatics:
% Application Notes (up to 2 pages; this is approx. 1300 words or 1000 words plus
% one figure) Applications Notes are short descriptions of novel software or new
% algorithm implementations, databases and network services (web servers, and
% interfaces). Software or data must be freely available to non-commercial users.
% Availability and Implementation must be clearly stated in the article. Authors
% must also ensure that the software is available for a full TWO YEARS following
% publication. Web services must not require mandatory registration by the user.
% Additional Supplementary data can be published online-only by the journal. This
% supplementary material should be referred to in the abstract of the Application
% Note. If describing software, the software should run under nearly all
% conditions on a wide range of machines. Web servers should not be browser
% specific. Application Notes must not describe trivial utilities, nor involve
% significant investment of time for the user to install. The name of the
% application should be included in the title.

\subtitle{Subject Section}

\title[Recording the pedigree]{ftprime: Recording the pedigree: efficient simulation of whole genomes}
\author{Corresponding Author\,$^{\text{\sfb 1,}*}$, Co-Author\,$^{\text{\sfb 2}}$ and Co-Author\,$^{\text{\sfb 2,}*}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{
    To use genomic data for inference and prediction it is often necessary to obtain whole-genome information
    from individual-based simulations,
    but the computational burden of tracking the genome of each simulated individual can be substantial.
    In this note we describe how to both (a) dramatically reduce this burden and 
    (b) efficiently record the entire history of the population.
    We do this by simulating only those loci that may affect reproduction (those having non-neutral variants),
    and recording the entire history of genetic inheritance in an efficient representation of the ancestral recombination graph,
    on which neutral mutations can be quickly placed afterwards.
    \plr{make more clear data structure was already developed? refer to 'tree sequence' by name?}
    The algorithm is implemented in python,
    and is designed to be easily used by any forwards-time simulation software.
\textbf{Availability and implementation:} \\
\textbf{Contact:} \\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}

\maketitle

Other title ideas: (I think it has to start with the `software name', for Bioinformatics applications notes anyhow)

Coalescence theory in forwards time

The ancestral recombination graph from individual-based simulations

History, and whole genomes too

\section{Introduction}

% OUTLINE
% 1. motivate need for whole-genome fwd-time simulations; point out that we only recently have the computing power to do this
% 2. explain ARG: explain that for forwards-time only need selected loci as by defn all others can be put on afterwards
% 3. back-of-the-envelope calculation to compare cost of tracking whole genomes versus putting mutations on ARG
% 4. review something about your efficient methods for storing/traversing tree sequence
% 5. write down algorithm used in ftprime to produce msprime-ready coalescence records
% 6. comparison of speed in really simple case (but whole-chromosome) using simupop in both cases

In recent years the increase in computing power --
gradual only in comparison to the increase in our ability to sequence genomes --
has made it possible to simulate the evolution of whole genomes of realistically-sized populations.
This major milestone allows researchers to investigate models of much greater complexity
than was previously acheivable.
Thus far, much of our understanding of how populations evolve
comes from models of one or a few linked loci \citep{ewens},
and/or results derived under the assumption of neutrality.
These are certainly important, but researchers today have widely differing views
on the practical importance of widespread selection \citep{encode,graur,backgroundsel}.
% Thus far what little analytical progress has been made understanding
% how potentially many functional genomic variants might evolve in ecologically realistic environments
% are hard to check due to a lack of realistic simulation methods.
Simulations of organisms carrying with functional, varying genomes
interacting with each other and spatially and temporally varying environments
will not only allow us to develop and test our understanding of evolution,
but also to produce quantitative predictions of eco-evolutionary dynamics --
for instance, the spread of insecticide resistance and behavior modification of mosquitos.

The most commonly used simulation methods to date rely on neutrality,
as this assumption coupled with random mating allows the use of coalescent methods \citep{wakeley}.
These drastically reduce the amount of required computation,
because they exploit a time-reversal duality of such models
to \emph{only} simulate the portion of history that determines the genomes in the modern population.
How does this work?
Suppose that to the \emph{population pedigree} --
the entire history of parent-offspring relationships of an entire population going back to a remote time --
we add information encoding the genetic outcomes of each ancestral meiosis --
who inherited which parts of which parental chromosomes.
This embellished graph is known as the \emph{ancestral recombination graph}, or ARG \citep{griffiths}.
Combined with ancestral genotypes and the origins of new mutations,
it completely specifies the genomic sequence of any individual in the population at any time.
However, much less than the entire ARG is needed to specify relationships between any given set of samples --
only those portions of it from which those samples have actually inherited, back to their most recent common ancestors.
For instance, \plr{add diagram of simple example?}
The assumptions of coalescent theory -- random mating and neutrality --
imply that the stochastic law of this random set forms a Markov process looked at \emph{backwards} in time,
and can hence be simulated without reference to the unnecessary remainder.
Although this point was in principle known since \citet{griffithsmarjoram},
only recent algorithmic advances made it possible to actually simulate the process correctly,
without approximation, across whole genomes \citep{msprime}.

These efficient methods are not available if either key assumption of coalescent theory --
neutrality or random mating -- are broken.
(Extensions can be made for only a few loci under selection \citep{hudson,krone}.)
Forwards-time, individual-based simulation, however, is burdened by recording and passing on entire genomes,
which typically have tens to hundreds of millions of varying sites.
Such simulations must explicitly record any non-neutral genetic variants,
because these affect the demographic process.
However, neutral variants, by definition, do not, 
and so if the demographic history is recorded
they can be simulated afterwards.

Wrap-up paragraph.  
This is what we do, which lets us both record history
and be more efficient.

\begin{methods}
\section{Methods and Implementation}

Since msprime provides a number of tools for manipulating tree sequences,
adding mutations,
and producing output (e.g., VCF files)
we need only determine a method for properly recording history and passing it to msprime.

\subsection{msprime: Tree sequences and coalescence records}

First we recall the data storage methods used by msprime \citep[see][for more details]{msprime}.

To record the history and identity of a set of sampled genomes,
it is only necessary to record commmon ancestry (or, \emph{coalescence}) events --
most recent common ancestors of distinct segments of sampled genomes --
and the identity and location of any mutations that these most recent common ancestors carry
that are not shared by any other recorded ancestor.
\plr{(not quite right)}

The topology of the ARG
is recorded by msprime as a list of Coalescence Records, which are tuples of the form
\texttt{(left,right,parent,children)},
which indicates that the individuals listed in \texttt{children}
have inherited the segment of genome between \texttt{left} and \texttt{right}
from the ancestor \texttt{parent}.
Individuals are identified by integers,
and there is also a table recording the \emph{birth time} of each individual.

For a set of coalescence records to be consistent,
it is required that:
\begin{enumerate}
    \item The birth time of any \texttt{children} must be strictly after the birth time of the \texttt{parent}.
    \item The set of intervals on which individual \texttt{a} is a child must be disjoint, for every \texttt{a}.
    \item The set of intervals on which individual \texttt{a} is a parent must be disjoint, for every \texttt{a}.
\end{enumerate}
These ensure that the coalescence records unambiguously specify a subset of the genealogy at any location on the genome,
but do not ensure that the entire history of the samples is specified.
Furthermore, to allow for efficient algorithms, msprime also requires that
\begin{enumerate} \setcounter{enumi}{3}
    \item The samples must be numbered \texttt{0,...,n-1}, and the smallest non-sampled label must be an internal node.
        \plr{perhaps remove this one}
    \item The list of \texttt{children} in a coalescence record must be sorted.
    \item Records must be sorted in nondecreasing order by birth time of the \texttt{parent}.
    \item Birth times must be strictly greater than zero.
\end{enumerate}

\subsection{Recording history}

Simulation software often gives a unique identifier to each \emph{diploid} individual,
but the natural exchangeable unit of consideration for the ARG (and msprime) is gamete --
so, each diploid individual is the union of two haploid individuals: 
one consisting of the set of chromosomes inherited maternally, the other paternally.
To translate between the two,
we adopt the convention that the maternally inherited set of chromosomes of diploid individual $k$
is given haploid ID $2k$, and the paternally inherited set $2k+1$.
ID numbers referred to below are haploid IDs.

Since we do not know from which ancestors
the eventual genetic samples will inherit,
to record the history of the samples
we must record the results of all reproduction events -- 
the entire ARG.
To do this, and conform with the format above, we maintain
an ordered dictionary \texttt{arg},
with one entry for each parent,
whose entries are of the form
\texttt{id:(time,records)},
where \texttt{id} is the integer ID of the parent,
\texttt{time} is the birth time of the parent,
and \texttt{records} is a list of nonoverlapping coalescence records, sorted by left endpoint.
We update this structure each time a new individual is born.
Suppose that individual $I$ is born at time $t$ to parents $p_0$ and $p_1$,
with recombination breakpoints $0=x_0 < x_1 < x_2 < \cdots < x_n \le L$,
so that $I$ inherits the genomic segment $[0,x_1)$ from $p_0$,
the segment $[x_1,x_2)$ from $p_1$, and so forth (here $n \ge 1$),
and where $L$ is the total length of the chromosome.
When this occurs, we:
\begin{enumerate}
    \item append a new entry \texttt{I:(t,[])} to the list \texttt{arg}; and
    \item for each $1 \le k \le n$, \texttt{merge} the new record \texttt{(x[k-1],x[k],p[k mod 2],(I,))}
        into the existing records in `arg[p[k mod 2]]`.
\end{enumerate}
The second step requires we merge the new record with existing ones
so as to retain the criteria described above.
To \texttt{merge} a record saying that individual \texttt{I} also inherits from this \texttt{parent} on interval \texttt{[a,b)}
into an existing set of records of the form \texttt{(left,right,parent,children)},
we need to 
\begin{enumerate}
    \item split any record for which \texttt{left<a<right} or \texttt{left<b<right} into two at the corresponding endpoint, 
    \item and then add \texttt{I} to \texttt{children} for any intervals overlapping \texttt{[a,b)}.
\end{enumerate}
This is done easily, maintaining the fact that records are not overlapping and are maintained in sorted order.
\plr{could alternatively write out algorithm but it is a bit nitpicky}

% To do this, to \texttt{merge} a record \texttt{(new.left,new.right,parent,new.children)}
% into an existing sorted list of \texttt{records},
% we do the following.
% \begin{enumerate} \setcounter{enumi}{-1}
%     \item Initialize \texttt{here=new.left} and \texttt{k=0}.
%     \item While \texttt{here<right}, let the next entry of \texttt{records} 
%         (ordered by left endpoint) be \texttt{(left,right,parent,children)}.
%         \begin{enumerate}
%             \item If \texttt{right<here}, go to the next record.
%             \item If \texttt{here<left}: 
%                 \begin{enumerate}
%                     \item insert a new record \texttt{(here,min(new.right,left),parent,new.children)} before the current record
%                     \item and set \texttt{here=min(new.right,left)}.
%                 \end{enumerate}
%             \item otherwise, if  TOO MANY STEPS
%         \end{enumerate}
% \end{enumerate}

\end{methods}
\section{Results and Discussion}

\section{Conclusion}

\section*{Acknowledgements}

\section*{Funding}

\bibliographystyle{natbib}
\bibliography{refs.bib}

\end{document}