### 2018-04-12 talks/poster/evo-wibo-2018-poster.tex
poster

\documentclass[fontscale=0.38,a0paper]{baposter}

\tracingstats=2

\usepackage{times}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{bm}
\usepackage{caption}
\usepackage[authoryear,round]{natbib}
\usepackage[hidelinks]{hyperref}

\usepackage{tabularx}
\usepackage{booktabs}

\usepackage{colortbl}

\usepackage{multicol}

%\usepackage{pgfbaselayers}
%\pgfdeclarelayer{background}
%\pgfdeclarelayer{foreground}
%\pgfsetlayers{background,main,foreground}

%\usepackage{tikz}
%\usetikzlibrary{arrows,backgrounds,patterns,matrix,shapes,fit,calc,shadows,plotmarks,decorations.pathmorphing,positioning,trees}

\renewcommand{\familydefault}{\sfdefault}
\usepackage{helvet}
\usepackage{bookman}
\usepackage{palatino}

% My stuff
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
% end my stuff

\selectcolormodel{cmyk}

% \graphicspath{{images/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newcommand{\sub}[2]{\ensuremath{#1_{\textrm{#2}}}}
%\newcommand{\gen}[2]{\ensuremath{\textrm{#1}_{\textrm{#2}}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{1em}
\setlength{\columnseprule}{0mm} % no vertical line

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Other Macros
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \renewcommand{\labelenumi}{\alph{enumi})}
% \renewcommand{\theenumi}{\alph{enumi})}
\newcommand{\postercaption}[1]{\begin{minipage}{\linewidth}\center\smaller
  {#1}\end{minipage}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Define some colors
\definecolor{silver}{cmyk}{0,0,0,0.3}
\definecolor{black}{cmyk}{0,0,0.0,1.0}
\definecolor{white}{cmyk}{0,0,0.0,0.0}
\definecolor{darkSilver}{cmyk}{0,0,0,0.1}

\definecolor{darkYellow}{cmyk}{0,0,1.0,0.5}
\definecolor{yellow}{cmyk}{0,0,0.9,0.0}
\definecolor{reddishyellow}{cmyk}{0,0.22,1.0,0.0}
\definecolor{lightyellow}{cmyk}{0,0,0.3,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}

\definecolor{DavisBlue}{cmyk}{1,0.56,0,0.34}
%    C = 100  M = 56  Y = 0 K = 34

\definecolor{HohenheimBlue}{cmyk}{1,0.5,0,0.45}
\definecolor{HohenheimLightLightBlue}{RGB}{243,250,255}
\definecolor{HohenheimLightBlue}{RGB}{215,221,235}
\definecolor{HohenheimDarkBlue}{RGB}{52,104,152}
\definecolor{darkCyan}{cmyk}{1,1,0,0.5}
\definecolor{cyan}{cmyk}{1,0,0,0.0}
\definecolor{lightcyan}{cmyk}{0.3,0,0,0.0}
\definecolor{lightercyan}{cmyk}{0.1,0,0,0.0}
\definecolor{lightestcyan}{cmyk}{0.05,0,0,0.0}

\typeout{Poster Starts}
%Define custom background
\background{
  \begin{tikzpicture}[remember picture,overlay]%
    \draw (current page.north west)+(-2em,2em) node[anchor=north west] {\includegraphics[height=\textheight]{silhouettes_background}};
  \end{tikzpicture}%
}

\newlength{\leftimgwidth}
\begin{poster}%
  % Poster Options
  {
  % Columns and Column spacing
  columns=3,
  colspacing=1.5em,
  % Background
  % background=user,
  background=shade-tb,
  % Poster header options
  eyecatcher=yes,
  posterheaderheight=0.15\textheight,
  % Format of textbox header
  headerborder=open,
  % headershape=small-rounded,
  headershape=roundedright,
  headershade=shade-tb,
  % headershade=plain,
  headerfont=\Large\textsf, %Sans Serif
  % Format of textbox
  boxborder=small-rounded,
  % boxborder=roundedleft,    
  boxshade=plain,
  linewidth=0.5pt,
  % Color style
  % bgColorOne=lightestcyan,
  % bgColorTwo=lightcyan,
  bgColorOne=white,
  bgColorTwo=white,
  % borderColor=cyan,
  borderColor=black,
  headerColorOne=DavisBlue,
  % headerColorOne=cyan,
  headerColorTwo=black,
  % headerFontColor=black,
  headerFontColor=white,
  % boxColorOne=lightercyan,
  boxColorOne=HohenheimLightLightBlue,
  boxColorTwo=white,
  % Show grid to help with alignment
  grid=no
  % grid=yes
  }
  % Eye Catcher
  {
  \makebox[15em][r]{%
    \begin{minipage}{15em}
       \hfill
       \includegraphics[width=15em]{tskit_logo} \\
    \end{minipage}
    }
  }
  % Title
  {\sf %Sans Serif
  %\bf% %bold  
  \vspace{0.5em}
     \textbf{\textcolor{DavisBlue}{Recording pedigrees for fast genome simulation and analysis}}\vspace{0.5em}}
  % Authors
  {\sf %Sans Serif
    Peter Ralph$^{\ddagger}$ \\  \vspace{-1.0mm}
    {\small \textit{$\ddagger$ Mathematics and Biology, University of Oregon} }\\
    {\textbf{code:}  \url{https://github.com/tskit-dev} }\\
    {\textbf{preprint:}  \url{https://www.biorxiv.org/content/early/2018/01/26/248500} }\\
  }
  % Project logo
  {
    \makebox[15em][r]{%
      \begin{minipage}{12em}
        \hfill
          \includegraphics[width=12em]{UOSignature-STK-BLK} \\
      \end{minipage}
    }
  }

  % Width of left inset image
%  \setlength{\leftimgwidth}{0.78em+0.0em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Tree sequences: all the genealogies}{name=intro,column=0,row=0,span=1}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For a set of sampled chromosomes,
at each position along the genome there is a genealogical tree
that says how they are related.

A \textbf{tree sequence} describes this sequence of trees.

\emph{Observations:}

\begin{itemize}

    \item The \emph{pedigree} (parental relationships) plus crossover locations
        would give us the tree sequence for \emph{everyone, ever}.

    \item Much less can fully describe the history relevant to a \emph{sample} of genomes.

    \item This information is equivalent to the Ancestral Recombination Graph (ARG).

\end{itemize}

\citep{kelleher2016efficient}
introduced the \textbf{tree sequence} data structure
for a fast coalescent simulator, \texttt{msprime}.

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Tables: a data structure for tree sequences}{name=tables,column=0,span=1,below=intro}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}

        \item \textbf{Edges:}   Who inherits from who; only \emph{necessary} for coalescent events.
            \emph{Records:} interval (left, right); parent node; child node.

        \item \textbf{Nodes:}   The ancestors those happen in.
            \emph{Records:} time ago (of birth); ID (implicit).

        \item \textbf{Mutations:}   When state changes along the tree.
            \emph{Records:} site it occurred at; node it occurred in; derived state.

        \item \textbf{Sites:}   Where mutations fall on the genome.
            \emph{Records:} genomic position; ancestral (root) state; ID (implicit).

    \end{itemize}

    \begin{center}
        \includegraphics[width=0.75\textwidth]{../file_size}
    \end{center}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Example}{name=ex,column=0,below=tables}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \includegraphics[width=\textwidth]{example_tree_sequence}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Another example}{name=ex2,column=0,below=ex}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \noindent
    \tiny
\hfill
\begin{tabular}{|rr|}
  \hline
    \textbf{Nodes:} & \\
id & time \\ 
  \hline
  0 & 0.00 \\ 
    1 & 0.00 \\ 
    2 & 0.00 \\ 
    3 & 0.00 \\ 
    4 & 0.00 \\ 
    5 & 0.00 \\ 
    6 & 0.00 \\ 
    7 & 0.00 \\ 
    8 & 0.09 \\ 
    9 & 0.31 \\ 
   10 & 0.39 \\ 
   11 & 0.41 \\ 
   12 & 0.43 \\ 
   13 & 0.97 \\ 
   14 & 1.04 \\ 
   15 & 1.21 \\ 
   16 & 1.36 \\ 
   17 & 1.45 \\ 
   18 & 2.41 \\ 
   19 & 2.46 \\ 
   20 & 2.85 \\ 
   21 & 3.84 \\ 
   22 & 4.45 \\ 
   \hline
\end{tabular}
\hfill
\begin{tabular}{|rrrr|}
  \hline
    \textbf{Edges:} & & & \\
left & right & parent & child \\ 
  \hline
0.00 & 1.00 &   8 &   2 \\ 
  0.00 & 1.00 &   8 &   6 \\ 
  0.00 & 1.00 &   9 &   5 \\ 
  0.00 & 1.00 &   9 &   7 \\ 
  0.00 & 1.00 &  10 &   3 \\ 
  0.00 & 1.00 &  10 &   8 \\ 
  0.00 & 1.00 &  11 &   0 \\ 
  0.00 & 1.00 &  11 &   1 \\ 
  0.00 & 1.00 &  12 &   9 \\ 
  0.00 & 1.00 &  12 &  10 \\ 
  0.99 & 1.00 &  13 &   4 \\ 
  0.99 & 1.00 &  13 &  12 \\ 
  0.00 & 0.19 &  14 &  11 \\ 
  0.00 & 0.19 &  14 &  12 \\ 
  0.93 & 0.99 &  15 &   4 \\ 
  0.93 & 1.00 &  15 &  11 \\ 
  0.99 & 1.00 &  15 &  13 \\ 
  0.30 & 0.54 &  16 &   4 \\ 
  0.91 & 0.93 &  16 &   4 \\ 
  0.30 & 0.54 &  16 &  11 \\ 
  0.91 & 0.93 &  16 &  11 \\ 
  0.54 & 0.69 &  17 &   4 \\ 
   \hline
\end{tabular}
\begin{tabular}{|rrrr|}
  \hline
    \textbf{cont'd} & ... &  ... & ... \\
left & right & parent & child \\ 
  \hline
  0.54 & 0.69 &  17 &  12 \\ 
  0.69 & 0.91 &  18 &   4 \\ 
  0.69 & 0.91 &  18 &  11 \\ 
  0.00 & 0.30 &  19 &   4 \\ 
  0.69 & 0.69 &  19 &  11 \\ 
  0.19 & 0.46 &  19 &  12 \\ 
  0.69 & 0.99 &  19 &  12 \\ 
  0.00 & 0.19 &  19 &  14 \\ 
  0.93 & 0.99 &  19 &  15 \\ 
  0.30 & 0.46 &  19 &  16 \\ 
  0.91 & 0.93 &  19 &  16 \\ 
  0.69 & 0.69 &  19 &  17 \\ 
  0.69 & 0.91 &  19 &  18 \\ 
  0.19 & 0.30 &  20 &  11 \\ 
  0.19 & 0.30 &  20 &  19 \\ 
  0.60 & 0.69 &  21 &  11 \\ 
  0.60 & 0.69 &  21 &  17 \\ 
  0.54 & 0.60 &  22 &  11 \\ 
  0.46 & 0.54 &  22 &  12 \\ 
  0.46 & 0.54 &  22 &  16 \\ 
  0.54 & 0.60 &  22 &  17 \\ 
   \hline
\end{tabular}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Contributions}{name=contrib,column=1,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}
        \compresslist
    \item Jerome Kelleher -- algorithms, data structures, and core development
    \item Kevin Thornton -- implementation, profiling with \texttt{fwdpp},
    \item Jaime Ashander -- implementation, profiling with \texttt{simuPOP}
    \item Ben Haller -- implementation in \texttt{SLiM}
    \item Jared Galloway -- implementation in \texttt{SLiM}
    \end{itemize}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Record history in forwards simulations?}{name=forwards,column=1,below=contrib}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}

        \item Coalescent simulations are \emph{much faster}
            than forwards-time, individual-based simulations
            because they don't have to keep track of \emph{everyone},
            only the ancestors of your sample.

        \item \textbf{But:} selection, or sufficient geographic structure,
            break the assumptions of coalescent theory.

        \item If we \emph{record the tree sequence}
            that relates everyone to everyone else,
            after the simulation is over we can put neutral mutations down on the trees.

        \item Since neutral mutations don't affect demography,
            this is \emph{equivalent} to having kept track of them throughout.

        \item This means recording the entire genetic history of \textbf{everyone} in the population, \textbf{ever}.

    \end{itemize}

\includegraphics[width=7em]{../finger_right} {\Large \emph{This may be a bad idea.}}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{How to do it}{name=howfwd,column=1,below=forwards}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{enumerate}

        \item add each gamete to the Node Table,
        \item add entries to the Edge Table
            recording which parent each gamete inherited each bit of genome from, and
        \item add any new selected mutations to the Mutation Table 
            and (if necessary) their locations to the Site Table.

    \end{enumerate}

\textbf{But,} we won't end up needing the entire history of everyone ever.

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Simplification}{name=simplify,column=1,below=howfwd}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    Given an input tree sequence
    and a subset of its nodes (the \emph{samples}),
    we want a new tree sequence for which:

    \begin{enumerate}
        \item All marginal trees match the corresponding subtree 
            in the input tree sequence.

        \item Every non-sample node in marginal trees has at least two children.

        \item All nodes and edges ancestral to at least one sample.

        \item No adjacent redundant edges \\
            (e.g., $(\ell, x, p, c) + (x, r, p, c) \rightarrow (\ell, r, p, c)$).
    \end{enumerate}

    To simplify a tree sequence
    to the history of the \emph{samples},
    we:

    \begin{itemize}
        \item Paint each \emph{sampled} chromosome a distinct color.

        \item Moving back up the tree sequence,
            copy colors of each chromosome to the parental chromosomes
            they inherited from.

        \item If two colors go in the same spot (\emph{coalescence}),
            replace with a new color (unique to that ancestor).
            Output a node for the ancestor and an edge for the coalescence.

        \item Once all colors have coalesced in a given segment,
            stop propagating it.
    \end{itemize}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Simulations, $50\times$ faster}{name=fast,column=2,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    Recording, simplifying, and output of tables, was done with \texttt{C} code in \texttt{tskit}.
    The simulation was done in \texttt{fwdpp} (C++ by Kevin Thornton),
    connected with \texttt{pybind11} and \texttt{numpy}.
    \textit{Machine:} Ubuntu / 2x 2.6 GHz Intel E5-2650 CPU.

    \emph{Simulation parameters:}
    \begin{itemize}
            \compresslist
        \item  Wright-Fisher population of size $N$, for $10N$ generations
        \item  neutral mutation rate $\mu$ equal to recombination rate $r$ per gamete
        \item  many, weakly deleterious mutations: rate $\mu/100$ with
            $s$ exponentially distributed with mean $2.5/N$.
    \end{itemize}

    \emph{Note:}
    simulations that recorded tree sequences ("pedigree recording")
    had neutral mutation rate was zero,
    but neutral mutations were added \emph{afterwards}.

    \begin{center}
    \includegraphics[width=\textwidth]{rawspeed}

    \includegraphics[width=0.75\textwidth]{speedup}
    \end{center}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Fast statistics computation}{name=stats,column=2,below=fast}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \textbf{A general class of statistics:}
    any polynomial $f(p_1, \ldots, p_k)$ gives a statistic on $k$ populations,
    by setting $p_i$ to the allele frequency in population $i$ and averaging $f$ across loci.
    \emph{(Example: divergence is $p_1(1-p-2) + p-2(1-p_1)$.)}

    The polynomial defines a weighting function on edges of the trees in the tree sequence,
    and the statistic can be calculated by summing across mutations
    weighted by the edge they occur on.
    \emph{(Example: proportion of pairs from the two populations on opposite sides of the edge.)}

    \textbf{Computation on a tree sequence:}
    If we compute weights on the first tree, then we only need update those weights above nodes that change
    in the second tree.

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{Summary}{name=summary,column=2,below=stats}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    Tree sequences 
    \begin{itemize}
        \item \ldots can make your simulations much faster
        \item \emph{because} they record genealogical history.
        \item (And, then you have genealogies!)
        \item \ldots can store genomic data \emph{very} efficiently,
        \item and compute with genomic data \emph{very} quickly.
    \end{itemize}

    \emph{Why} are they so useful?
    It is a compression scheme, provided by the process itself that produced the data.
    \emph{(think: optimal compression into shared haplotypes)}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{}{name=references,column=2,below=summary}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \scriptsize
  \renewcommand{\section}[2]{\vskip 0.0em}
  \bibliographystyle{abbrvnat}
  \setlength{\bibsep}{0.0pt}
  \bibliography{../../references}
  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\headerbox{}{name=trees,column=0,above=bottom,span=3}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \includegraphics[width=0.085\textwidth]{tree0}
    \includegraphics[width=0.085\textwidth]{tree1}
    \includegraphics[width=0.085\textwidth]{tree2}
    \includegraphics[width=0.085\textwidth]{tree3}
    \includegraphics[width=0.085\textwidth]{tree4}
    \includegraphics[width=0.085\textwidth]{tree5}
    \includegraphics[width=0.085\textwidth]{tree6}
    \includegraphics[width=0.085\textwidth]{tree7}
    \includegraphics[width=0.085\textwidth]{tree8}
    \includegraphics[width=0.085\textwidth]{tree9}
    \includegraphics[width=0.085\textwidth]{tree10}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{poster}%
\end{document}